---
title: "Results"
author: "Waldemar Ortiz-Calo"
date: "1/3/2021"
output: html_document
---
```{r readData,echo=FALSE}
# Library
library(ggplot2)
library(magrittr)

# Data 

# General Data
HomeRangeSamples <- read.csv("D:\\Drive\\Research\\RECaP\\2.INPROGRESS\\1.SpatialDimensionality\\3.Output\\HomeRangeSamples.csv") %>% subset(TRI > 0)

# Outlier Removal
GenOutlier <- subset(HomeRangeSamples, PercentDifference < 30) %>% subset(TRI > 0)
```

These are the results of my simulations. Prior to cleaning the results, there was a total of 50,000 home range samples. This was a representative sample of 2500 raster tiles (10.6 %) with 12,500 home range samples for each size category (5, 10, 100, 250 km^2^). 

Once the data was collected, I then cleaned the data by removing any NAs and removing any samples that had an average TRI of 0. I did this because a TRI of zero represents that there was no elevation data within that home range. Once I did this, I had a total of 40,972 samples. Furthermore, I eliminated outlier values by discarding samples that had extremely high values. This resulted in a final data set with 39,513 samples. 


### I. Percent Difference vs TRI  


```{r,eval = T,echo=FALSE,message = F}

# General Plot 
ggplot(HomeRangeSamples, aes(TRI,PercentDifference))+
  geom_point()+
  ggtitle("Percent Difference v. TRI")+
  theme_bw()+
  geom_smooth()

# Faceted Plot 
ggplot(HomeRangeSamples, aes(TRI,PercentDifference))+
  geom_point()+
  ggtitle("Percent Difference v. TRI")+
  theme_bw()+
  geom_smooth()+
  facet_grid(. ~ factor(Size_Category,levels = c("5km","10km","100km","250km")))

```

### II. Percent Difference vs TRI (Without Outliers) 


```{r,eval = T,echo=FALSE,message = F}

#     Percent Difference v TRI (Outlier Removal)                            ####

# General Plot
ggplot(GenOutlier, aes(TRI,PercentDifference))+
  geom_point()+
  ggtitle("Percent Difference v. TRI")+
  theme_bw()+
  geom_smooth()

# Faceted Plot 
ggplot(GenOutlier, aes(TRI,PercentDifference))+
  geom_point()+
  ggtitle("Percent Difference v. TRI")+
  theme_bw()+
  geom_smooth()+
  facet_grid(. ~ factor(Size_Category,levels = c("5km","10km","100km","250km")))



```

### III. Distribution of Percent Difference 

```{r,eval = T,echo=FALSE,message = F,warning=F}

# Distribution General
ggplot(GenOutlier) + geom_histogram(aes(PercentDifference))+
  ggtitle("Percent Difference Distribution")+
  xlab("Percent Difference")+
  theme_bw()
  
# Distribution Faceted
ggplot(GenOutlier) + geom_histogram(aes(GenOutlier$PercentDifference))+
  ggtitle("Percent Difference Distribution")+
  xlab("Percent Difference")+
  theme_bw()+
  facet_grid(. ~ factor(Size_Category,levels = c("5km","10km","100km","250km")))

```
